Inducing Semantic Frames on a Very Large Corpus of Syntactic-Ngrams

### 2015-01-05 Paper/paperN.tex
spell check, abstract

\begin{abstract}
The abstract paragraph should be indented 1/2~inch (3~picas) on both left and
right-hand margins. Use 10~point type, with a vertical spacing of 11~points.
The word \textbf{Abstract} must be centered, bold, and in point size 12. Two
line spaces precede the abstract. The abstract must be limited to one
paragraph.
\end{abstract}

\section{Introduction}
The growing body of digital publications contains too much information for a single human being to process; it is even hard to keep up to date on a particular topic. It would be useful to be aided by technology that creates reliable summarizations. As a natural language processing task, summarization consists of two parts: first, extract the semantic content of a text, and then represent it in a condensed human readable output. The first part of such a technology, extracting semantic content, is what we deal with in this project. This step may also be useful on its own, for example, in automatic stock trading systems. Knowing what is published about publicly traded companies at the moment of publication can increase prophet margins for so-called high-frequnecy trading algorithms. Semantic representations have applications in other natural language procesing as well. For example, interlingua approaches to machine translation try to match sentences in the sorce and target language to a common generic semantic representation \cite{boas2005}.

Semantic content of a sentence can be represented in a frame. The main verbphrase of the sentence is the head of the frame and can be interpreted as predicate, representing the core of the episode described by the sentence. The trick is to find out what the arguments of the predicate are. In linguistics, argument selection is called case grammar \citep{dowty1991} and arguments (or roles) have specific names as \textit{agent - patient}. In computational systems names are not important and can be asigned when frame induction is complete.

The problem of semantic content can be illustrated by this simple example: \textit{'Alice buys a bread from the baker'} descibes the exact same situation as \textit{`The baker sells Alice a bread'}. Thus the semantic content of these sentences can be represented in one frame, where the head of the frame represents the core of the episode or situation and involvement is expressed by assigning roles to things or people. In general these roles can be anything like location, time, manner etc. The frame for the sentences above is depicted in table \ref{framex}.
\begin{wraptable}{r}{.3\textwidth}
  \centering
 % \begin{table}
\begin{tabular}{|l l|}
  \hline
  \textit{\small predicate:\normalsize}&Transaction\\
  \hline
  \hline
  \textit{\small role 1:\normalsize} &baker\\
  \textit{\small role 2:\normalsize} &Alice\\
  \textit{\small role 3:\normalsize} &bread\\
  \hline
\end{tabular}
%\end{table}
\caption{Simple frame}
\label{framex}
\end{wraptable}
As the example above shows, the roles are represented by different syntactic parts of speech in both sentences: the object and subject are switched. Furthermore, the verbs are different. Even given the knowledge of specific verbs that bear the same meaning with switched subject/indirect-object, we need to generate semantic representations on top of syntax in order to let machines (act as if they) understand language and process new text in a meaningfull way. Hence, automatic semantic analysis and representation is required for further automatization.

@@

In this project we investigate whether probabalistic latent-variable models can cluster verbs based on their meaning, which is inferred from the arguments with wich it occurs. These clusters then make up a broader, but hopefully distinct, meaning. The data we use contains triples of verb-subject-object (see section \ref{data}), so induced frames contain only two roles. The models (described in section \ref{models}) induce frames based on the assumption the meaning of a verb relative to meaning of other verbs can be induced from its context. Hence the broader meaning is latent. To evaluate our models we use two metrics, one of which uses framenet data to check wheter the induced frames are similar to clusters of verbs in framenet, the other to check wheter the resulting frames are coherent, i.e. wheter the found clusters are not completely random. See section \ref{results} for more detail.

%\section{Related Work}
We based our work on two papers: \textit{`Inducing a semantically annotated lexicon via em-based clustering.'} by \citeauthor{rooth1999} and \textit{`Learning frames from text with an unsupervised latent variable model.'} by \citeauthor{oconnor2013}. In the first article clustering is realized by applying expectation maximalization on a simple cooccurance model. This model produces frames in the form of two distributions, one for the predicate, a distribution over verbs, and one for the arguments, a nouns. Following O'Connor, we modify this model slightly to accomodate verb-subject-object triples. A triple is assigned to the frame for which is probability of occuring is highest. We will use this model as our baseline (as in the O'Connor paper). The second paper covers two additional models, of which we implemented one. This model differs from the baseline model by introducing the assumption that a document generates sentences frome a restricted set of frames. This assumption is justified (as in latent Dirichelt allocation) \cite{blei2003} by the intuition that documents generally have a limited number of topics. In our data the tuples are not annotated with documents so we used this assumption in such a way that each verb covers one document.

\section{Induction Models}
\label{models}
We consider two probabalistic latent-variable models that learn semantic frames
from verb-subject-object tripples (VSO's).
The goal of these models is to use the syntactic information given by a VSO to
infer what kind of event it describes (i.e., to which semantic frame it belongs).

@@

    \begin{snugshade}
    \scriptsize
    For each frame $f=1..F$:\\
    \hspace*{15pt} For each argument $a\in\{v,s,o\}$:\\
    \hspace*{30pt} Draw a distribution $\phi_f^a\sim Dirichlet(\beta)$\\
    For each document $d=1..D$:\\
    \hspace*{15pt} Draw a distrbution $\theta_d \sim Dirichlet(\alpha)$\\
    \hspace*{30pt} For each $i = 1..N^d$:\\
    \hspace*{30pt} Draw a frame $f \sim \theta^d$.\\
    \hspace*{30pt} Draw a verb $v \sim \phi_f^v$\\
    \hspace*{30pt} Draw a subject $s \sim \phi_f^s$\\
    \hspace*{30pt} Draw an object $o \sim \phi_f^o$
    \end{snugshade}

@@

This approach models each VSO independenly, without considering any further context.
Word distributions are completely independent between arguments and accross frames.
Tuples are clustered according to which frame's three distributions best fit the 
three arguments; that is,

@@

Model 0 is based on the apporach originally proposed in \citet{rooth1999}, which 
considers noun-verb pairs. Following \citet{oconnor2013}, we expand the model
to use the limited syntax of VSOs.

@@

It is in the nature of documents that they have some degree of internal semantic
coherence. Put more plainly, a document is usually \emph{about} something. We
know from topic modeling that LDA with Gibbs sampling can be very successful at
classifying documents according to some semantic features \citep{blei2003}.
Model 1 leverages this fact by making the assmumption that a given document is 
likely to have a sparse distribution of frames. As in LDA, the document-level
sparsity is enforced by a dirichlet prior on frame distributions (see figure
\ref{gen1} for details). This assumption breaks the independence between tuples 
that is maintained by Model 0 is dropped in Model 1.

@@

\begin{align*}
 P(\mathbf{w},\mathbf{f}|\alpha,\beta)
&=\int_\phi\int_\theta P(\mathbf{w},\mathbf{f}|\theta,\phi,\alpha,\beta)\\
&=     \int_\phi P(\mathbf{w}|\mathbf{f},\phi )\,P(\phi|\beta)
\times \int_\theta P(\mathbf{f}|\theta)\,P(\theta|\alpha)\\
\intertext{These integrals are both dirichlet multinomials:}
&=\frac{DCM_i^{frames}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{frames}(\mathbf{f}_{-ij})}\times
\prod_a^{\{v,s,o\}}\frac{DCM_i^{vocab^a}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{vocab^a}(\mathbf{f}_{-ij})}
\end{align*}

We prefer not to consider the subject and object vocabularies independently 
(i.e., $V^s = V^o$). This is more in line with the theory of semantic frames 
since in principle, a noun playing a particlar semantic role may appear in either 
syntactic position.

@@

The data used in this probject (see section \ref{data}) doesn't contain any document-level
information. As a result, we adapt model 1 by grouping VSOs with the same verb into ``documents''.
This move leverages the dirichlet prior $\alpha$ to enforce sparcity of frames 
accross verbs, but of course cannot capture the information given by the kind of
document-level semantic sparsity seen in LDA.\footnote{When we use verbs as
documents (i.e, $D = V^v$), since $w_{ij}^v$ is the same for all verbs in document $d_i$ 
(and $d_i$ is the only document with that verb), it follows that the number of VSO's in document 
$d_i$ belonging to frame $f$ is the same as the number of VSO's with $w_{ij}^v$ as their verb assigned to 
frame $f$; that is, $\tilde c(d_i, f) = \tilde c(f, w_{ij}^v)$.}

@@

A syntactic-ngram is a $k$-word rooted subtree of some sentence.
Google syntactic-ngrams is a corpus of ngrams extracted from 3.5 million English language 
books and labled with Penn Treebank part-of-speech tags Stanford-basic dependency 
lables \citep{ngrams2013}.
Each ngram also includes frequency counts by year.

@@

\section{Experiments}
The number of frames in both models must be determined before induction can take place. When choosing for a low number of frames the clusters will be bigger and thus the meaning of a cluster more vague. In contrast, a high number of frames bears the danger of overfitting the data, i.e. when unrelated verbs occur with the same arguments often they will be clustered and may even make up one cluster, while with a lower number of frames other verbs would be included, diminishing one of the unrelated verbs as noise. Following \citet{rooth1999}, we estimate that the appropriate number of frames to induce is at most a couple hundred. This assumption is based on philosophical arguments of Jerry Fodor, who proposes that there should be a very limited degree of genuine interdefinability between lexical items (here frames), and that the magnitude of the lexicon should therefore be approximately equal to the number of basic concepts \citep{fodor1998}.

@@

The $\alpha$ and $\beta$ hyperparameters in model 1 are known  to be hard to determine \citep{oconnor2013}, and therefore need to be experimentally optimized for our specific problem. We used an expensive, but easy to implement grid search to determine the best values for $\alpha$ and $\beta$. The grid search builds, for each posible parameter set, a model which is evaluated on the cross validation set. The highest scoring model is tested with the test set. We ran grid search for model 0 on $50$ - $200$ frames, with steps of $50$, and $\alpha$ of $0.1, 0.5, 1, 1.5$, and $2$. The grid search for Model 1 was the the same, additionally testing $\beta$ accross the same values as $\alpha$
\section{Results}
\label{results}

\subsection{Frame Coherency}
Verb clusters are based on the arguments with which they occuring in the data. So to test how well the models perform on clustering, the probability that the model generates a triple from the testset, should be greater than this triple with a random verb in place of the real one. This metric is the same as the one used by \citeauthor{rooth1999}. More formally: for a datapoint $(v,s,o)$ and a tuple $(v^r,s,o)$ where $v^r$ is a random chosen verb: $P(v\mid s,o) \geq P(v^r\mid s,o)$.
The results are percentages of datapoints which satisfy this condition. Note though that if the random verb-triple is assigned the same frame as its `parent' real triple, this metric loses its meaning, but since the number of frames is high this probabilty is low. Furthermore, one can assume that the occurences of same frame assignement cancel eachother out, i.e. the random triple has higher probabilty than the real one as many times as vice versa.

@@

\section{Future Work}
One big weakness of this project over the results of \citet{oconnor2013} is that 
we could not used document-level context to infer semantic information.
A possible direction of future reseach is to create models that take better 
advantage of the syntax-level context provided by syntactic ngrams. 
One easy way to do this would be to modify model 1 so that the argument 
vocabulary ranges over subject/object pairs instead of individual nouns.
or object.
More advanced models might take better advantage of the various kinds of arguments
in the verbargs dataset and their relation to one another.
For example, you could be gleen some semantic information about the relationship between
verbs and prepositional objects by looking at the preposition that mediates them: 
That you ''swim \emph{in} a lake'' versus ``ice skate \emph{on} a lake'' 
indicates that the relationship between swimming and lakes is different from the one between
ice skating and lakes.

>>>>

\begin{abstract}
We explore two models for the unsupervised induction of semantic frames on a 
very large corpus of verb-subject-object triples. Our work is bade on previous 
research that has had success on smaller, more focused data sets. By encouraging 
verb sparsity with a Dirichlet prior, one can improve on the results of our 
baseline model. We give some analysis of the frames that the improved model 
discovers, and suggest areas for further development.
\end{abstract}

\section{Introduction}
The growing body of digital publications contains too much information for a single human being to process; it is even hard to keep up to date on a particular topic. It would be useful to be aided by technology that creates reliable summarizations. As a natural language processing task, summarization consists of two parts: first, extract the semantic content of a text, and then represent it in a condensed human readable output. The first part of such a technology, extracting semantic content, is what we deal with in this project. This step may also be useful on its own, for example, in automatic stock trading systems. Knowing what is published about publicly traded companies at the moment of publication can increase prophet margins for so-called high-frequency trading algorithms. Semantic representations have applications in other natural language processing as well. For example, interlingua approaches to machine translation try to match sentences in the source and target language to a common generic semantic representation \cite{boas2005}.

Semantic content of a sentence can be represented in a frame. The main verb phrase of the sentence is the head of the frame and can be interpreted as predicate, representing the core of the episode described by the sentence. The trick is to find out what the arguments of the predicate are. In linguistics, argument selection is called case grammar \citep{dowty1991} and arguments (or roles) have specific names as \textit{agent - patient}. In computational systems names are not important and can be assigned when frame induction is complete.

The problem of semantic content can be illustrated by this simple example: \textit{'Alice buys a bread from the baker'} describes the exact same situation as \textit{`The baker sells Alice a bread'}. Thus the semantic content of these sentences can be represented in one frame, where the head of the frame represents the core of the episode or situation and involvement is expressed by assigning roles to things or people. In general these roles can be anything like location, time, manner etc. The frame for the sentences above is depicted in table \ref{framex}.
\begin{wraptable}{r}{.3\textwidth}
  \centering
 % \begin{table}
\begin{tabular}{|l l|}
  \hline
  \textit{\small predicate:\normalsize}&Transaction\\
  \hline
  \hline
  \textit{\small role 1:\normalsize} &baker\\
  \textit{\small role 2:\normalsize} &Alice\\
  \textit{\small role 3:\normalsize} &bread\\
  \hline
\end{tabular}
%\end{table}
\caption{Simple frame}
\label{framex}
\end{wraptable}
As the example above shows, the roles are represented by different syntactic parts of speech in both sentences: the object and subject are switched. Furthermore, the verbs are different. Even given the knowledge of specific verbs that bear the same meaning with switched subject/indirect-object, we need to generate semantic representations on top of syntax in order to let machines (act as if they) understand language and process new text in a meaningful way. Hence, automatic semantic analysis and representation is required for further automatization.

@@

In this project we investigate whether probabilistic latent-variable models can cluster verbs based on their meaning, which is inferred from the arguments with which it occurs. These clusters then make up a broader, but hopefully distinct, meaning. The data we use contains triples of verb-subject-object (see section \ref{data}), so induced frames contain only two roles. The models (described in section \ref{models}) induce frames based on the assumption the meaning of a verb relative to meaning of other verbs can be induced from its context. Hence the broader meaning is latent. To evaluate our models we use two metrics, one of which uses framenet data to check whether the induced frames are similar to clusters of verbs in framenet, the other to check whether the resulting frames are coherent, i.e. whether the found clusters are not completely random. See section \ref{results} for more detail.

%\section{Related Work}
We based our work on two papers: \textit{`Inducing a semantically annotated lexicon via em-based clustering.'} by \citeauthor{rooth1999} and \textit{`Learning frames from text with an unsupervised latent variable model.'} by \citeauthor{oconnor2013}. In the first article clustering is realized by applying expectation maximization on a simple cooccurance model. This model produces frames in the form of two distributions, one for the predicate, a distribution over verbs, and one for the arguments, a nouns. Following O'Connor, we modify this model slightly to accommodate verb-subject-object triples. A triple is assigned to the frame for which is probability of occurring is highest. We will use this model as our baseline (as in the O'Connor paper). The second paper covers two additional models, of which we implemented one. This model differs from the baseline model by introducing the assumption that a document generates sentences frame a restricted set of frames. This assumption is justified (as in latent Dirichlet allocation) \cite{blei2003} by the intuition that documents generally have a limited number of topics. In our data the tuples are not annotated with documents so we used this assumption in such a way that each verb covers one document.

\section{Induction Models}
\label{models}
We consider two probabilistic latent-variable models that learn semantic frames
from verb-subject-object triples (VSO's).
The goal of these models is to use the syntactic information given by a VSO to
infer what kind of event it describes (i.e., to which semantic frame it belongs).

@@

    \begin{snugshade}
    \scriptsize
    For each frame $f=1..F$:\\
    \hspace*{15pt} For each argument $a\in\{v,s,o\}$:\\
    \hspace*{30pt} Draw a distribution $\phi_f^a\sim Dirichlet(\beta)$\\
    For each document $d=1..D$:\\
    \hspace*{15pt} Draw a distribution $\theta_d \sim Dirichlet(\alpha)$\\
    \hspace*{30pt} For each $i = 1..N^d$:\\
    \hspace*{30pt} Draw a frame $f \sim \theta^d$.\\
    \hspace*{30pt} Draw a verb $v \sim \phi_f^v$\\
    \hspace*{30pt} Draw a subject $s \sim \phi_f^s$\\
    \hspace*{30pt} Draw an object $o \sim \phi_f^o$
    \end{snugshade}

@@

This approach models each VSO independently, without considering any further context.
Word distributions are completely independent between arguments and across frames.
Tuples are clustered according to which frame's three distributions best fit the 
three arguments; that is,

@@

Model 0 is based on the approach originally proposed in \citet{rooth1999}, which 
considers noun-verb pairs. Following \citet{oconnor2013}, we expand the model
to use the limited syntax of VSOs.

@@

It is in the nature of documents that they have some degree of internal semantic
coherence. Put more plainly, a document is usually \emph{about} something. We
know from topic modeling that LDA with Gibbs sampling can be very successful at
classifying documents according to some semantic features \citep{blei2003}.
Model 1 leverages this fact by making the assumption that a given document is 
likely to have a sparse distribution of frames. As in LDA, the document-level
sparsity is enforced by a Dirichlet prior on frame distributions (see figure
\ref{gen1} for details). This assumption breaks the independence between tuples 
that is maintained by Model 0 is dropped in Model 1.

@@

\begin{align*}
 P(\mathbf{w},\mathbf{f}|\alpha,\beta)
&=\int_\phi\int_\theta P(\mathbf{w},\mathbf{f}|\theta,\phi,\alpha,\beta)\\
&=     \int_\phi P(\mathbf{w}|\mathbf{f},\phi )\,P(\phi|\beta)
\times \int_\theta P(\mathbf{f}|\theta)\,P(\theta|\alpha)\\
\intertext{These integrals are both Dirichlet multinomials:}
&=\frac{DCM_i^{frames}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{frames}(\mathbf{f}_{-ij})}\times
\prod_a^{\{v,s,o\}}\frac{DCM_i^{vocab^a}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{vocab^a}(\mathbf{f}_{-ij})}
\end{align*}

We prefer not to consider the subject and object vocabularies independently 
(i.e., $V^s = V^o$). This is more in line with the theory of semantic frames 
since in principle, a noun playing a particular semantic role may appear in either 
syntactic position.

@@

The data used in this project (see section \ref{data}) doesn't contain any document-level
information. As a result, we adapt model 1 by grouping VSOs with the same verb into ``documents''.
This move leverages the Dirichlet prior $\alpha$ to enforce sparsity of frames 
across verbs, but of course cannot capture the information given by the kind of
document-level semantic sparsity seen in LDA.\footnote{When we use verbs as
documents (i.e, $D = V^v$), since $w_{ij}^v$ is the same for all verbs in document $d_i$ 
(and $d_i$ is the only document with that verb), it follows that the number of VSO's in document 
$d_i$ belonging to frame $f$ is the same as the number of VSO's with $w_{ij}^v$ as their verb assigned to 
frame $f$; that is, $\tilde c(d_i, f) = \tilde c(f, w_{ij}^v)$.}

@@

A syntactic-ngram is a $k$-word rooted subtree of some sentence.
Google syntactic-ngrams is a corpus of ngrams extracted from 3.5 million English language 
books and labled with Penn Treebank part-of-speech tags Stanford-basic dependency 
labels \citep{ngrams2013}.
Each ngram also includes frequency counts by year.

@@

\section{Experiments}
The number of frames in both models must be determined before induction can take place. When choosing for a low number of frames the clusters will be bigger and thus the meaning of a cluster more vague. In contrast, a high number of frames bears the danger of over-fitting the data, i.e. when unrelated verbs occur with the same arguments often they will be clustered and may even make up one cluster, while with a lower number of frames other verbs would be included, diminishing one of the unrelated verbs as noise. Following \citet{rooth1999}, we estimate that the appropriate number of frames to induce is at most a couple hundred. This assumption is based on philosophical arguments of Jerry Fodor, who proposes that there should be a very limited degree of genuine interdefinability between lexical items (here frames), and that the magnitude of the lexicon should therefore be approximately equal to the number of basic concepts \citep{fodor1998}.

@@

The $\alpha$ and $\beta$ hyperparameters in model 1 are known  to be hard to determine \citep{oconnor2013}, and therefore need to be experimentally optimized for our specific problem. We used an expensive, but easy to implement grid search to determine the best values for $\alpha$ and $\beta$. The grid search builds, for each posible parameter set, a model which is evaluated on the cross validation set. The highest scoring model is tested with the test set. We ran grid search for model 0 on $50$ - $200$ frames, with steps of $50$, and $\alpha$ of $0.1, 0.5, 1, 1.5$, and $2$. The grid search for Model 1 was the the same, additionally testing $\beta$ across the same values as $\alpha$
\section{Results}
\label{results}

\subsection{Frame Coherency}
Verb clusters are based on the arguments with which they occurring in the data. So to test how well the models perform on clustering, the probability that the model generates a triple from the testset, should be greater than this triple with a random verb in place of the real one. This metric is the same as the one used by \citeauthor{rooth1999}. More formally: for a datapoint $(v,s,o)$ and a tuple $(v^r,s,o)$ where $v^r$ is a random chosen verb: $P(v\mid s,o) \geq P(v^r\mid s,o)$.
The results are percentages of data points which satisfy this condition. Note though that if the random verb-triple is assigned the same frame as its `parent' real triple, this metric loses its meaning, but since the number of frames is high this probability is low. Furthermore, one can assume that the occurrences of same frame assignment cancel each other out, i.e. the random triple has higher probability than the real one as many times as vice versa.

@@

\section{Future Work}
One big weakness of this project over the results of \citet{oconnor2013} is that 
we could not used document-level context to infer semantic information.
A possible direction of future research is to create models that take better 
advantage of the syntax-level context provided by syntactic ngrams. 
One easy way to do this would be to modify model 1 so that the argument 
vocabulary ranges over subject/object pairs instead of individual nouns.
or object.
More advanced models might take better advantage of the various kinds of arguments
in the verbargs dataset and their relation to one another.
For example, you could be glean some semantic information about the relationship between
verbs and prepositional objects by looking at the preposition that mediates them: 
That you ''swim \emph{in} a lake'' versus ``ice skate \emph{on} a lake'' 
indicates that the relationship between swimming and lakes is different from the one between
ice skating and lakes.
